---
output:
  html_document: default
  pdf_document: default
---
Courser Reproducible Research Week 2 Project
============================================

First, let's read the data. And then let's look at the first 20 observations

```{r}
dataset <- read.csv('activity.csv')
head(dataset,20)
```

Ok, how much data we have?

```{r}
dim(dataset)
```

Fine, how does the data look like

```{r}
summary(dataset)
str(dataset)
```

Great, let's explore it then!

Let's see a histogram of daily steps. We will need to:
- Aggregate the 'steps' column to the 'date' column so we have total steps per day
- Afterwards, we will plot a histogram with BaseR

```{r}
dataset_by_day <- aggregate(dataset$steps, by=list(dataset$date), sum)
hist(dataset_by_day$x, 
        xlab = 'Steps Per Day', 
        main = 'Histogram of Steps Per Day', 
        ylab = 'Days')
```

Cool, now let's compute the mean and median of steps taken per day

```{r}
mean(dataset_by_day$x, na.rm = T)
median(dataset_by_day$x, na.rm = T)
```

Now, we would like to see a time-series of the average steps taken per day.
First we create another dataset with average number of steps per day, then plot!

```{r}
dataset_avg_by_day <- aggregate(dataset$steps, by=list(dataset$date), mean)
dataset_avg_by_day$Group.1 <- as.Date(dataset_avg_by_day$Group.1)
plot(dataset_avg_by_day, 
     xlab='Time', 
     ylab='Average Steps', 
     main='Average Steps Per Day')
lines(dataset_avg_by_day)
```

Let's find the 5-minute interval that has the maximum number of steps

```{r}
#we first find the maximum number of steps and store it in max_steps
max_steps <- max(dataset$steps, na.rm = T)
#then use this value to find the interval
max_interval <- dataset$interval[dataset$steps==max_steps]
sum(max_interval, na.rm = T)
```

There are lots of missing value. It is a good idea to replace missing values by the mean

```{r}
dataset_imputed <- dataset
dataset_imputed$steps[is.na(dataset_imputed$steps)] <- mean(dataset$steps, na.rm = T)
summary(dataset_imputed)
```

Now, let's re-visit the histogram
```{r}
dataset_imputed_by_day <- aggregate(dataset_imputed$steps, by=list(dataset_imputed$date), sum)
hist(dataset_imputed_by_day$x, 
        xlab = 'Steps Per Day', 
        main = 'Histogram of Steps Per Day', 
        ylab = 'Days')
```

Let's make a time-series for the steps once for weekdays and once for weekends

```{r}
dataset_with_day_names <- dataset_imputed
dataset_with_day_names$dayname <- weekdays(as.Date(dataset_with_day_names$date))
dataset_weekends <- dataset_with_day_names[dataset_with_day_names$dayname == 'Friday' | 
                                             dataset_with_day_names$dayname == 'Saturday',]
dataset_weekdays <- dataset_with_day_names[dataset_with_day_names$dayname == 'Sunday'| 
                                           dataset_with_day_names$dayname == 'Monday' |
                                             dataset_with_day_names$dayname == 'Tuesday'|
                                             dataset_with_day_names$dayname == 'Wednesday'|
                                             dataset_with_day_names$dayname == 'Thursday', ]
dim(dataset_weekdays)
dim(dataset_weekends)
head(dataset_weekdays)
head(dataset_weekends)
dataset_weekdays_avg <- aggregate(dataset_weekdays$steps, by=list(dataset_weekdays$date), mean)
dataset_weekends_avg <- aggregate(dataset_weekends$step, by=list(dataset_weekends$date), mean)
head(dataset_weekdays_avg)
head(dataset_weekends_avg)
dataset_weekdays_avg$Group.1 <- as.Date(dataset_weekdays_avg$Group.1)
dataset_weekends_avg$Group.1 <- as.Date(dataset_weekends_avg$Group.1)
```

That was lots of preparing!
Now the plot

```{r fig.height=10}
par(mfrow=c(2,1))
plot(dataset_weekdays_avg, xlab = 'Days', ylab = 'Average Steps',
     main = 'Average Steps in Weekdays')
lines(dataset_weekdays_avg)
plot(dataset_weekends_avg, xlab = 'Days', ylab = 'Average Steps',
     main = 'Average Steps in Weekends')
lines(dataset_weekends_avg)
```
